---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: yes
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Load the dataset and convert data type to Date for date column for further analysis.
```{r, echo=TRUE}
act <- read.csv("activity.csv")
act$date <- as.Date(act$date)
```

## What is mean total number of steps taken per day?

Histogram of the total number of steps taken each day
```{r, echo=TRUE}
plot(aggregate(steps ~ date, data = act, sum, na.rm = TRUE), type = "h", main = "Total Number of steps per Day")
```

Following is the mean and median values of the total number of steps.
```{r, echo=TRUE}
mean(act$steps,na.rm = TRUE)
median(act$steps,na.rm = TRUE)
```

And, following is showing each mean and median stpes values per each day.
```{r, echo=TRUE}
aggregate(steps ~ date, data = act, mean, na.rm = TRUE)
aggregate(steps ~ date, data = act, median, na.rm = TRUE)
```

## What is the average daily activity pattern?

Time series plot of the 5 minutes interval and the average number of steps taken, averaged across all days.
```{r, echo=TRUE}
plot(aggregate(steps ~ interval, data = act, mean, na.rm = TRUE), type = "l", main = "Average Steps per Intervals")
```

Following is showing what interval contains the maximum number of steps.
```{r, echo=TRUE}
aggact <- aggregate(steps ~ interval, data = act, mean, na.rm = TRUE)
subset(aggact, aggact$steps == max(aggact[,2]))
```

## Imputing missing values

How many missing values?
```{r, echo=TRUE}
sum(is.na(act$steps))
```

The missing values are filling with the mean values per interval across all days.
```{r, echo=TRUE}
act$steps[is.na(act$steps)] <- aggact$steps[match(act$interval,aggact$interval)][which(is.na(act$steps))]

```

Histogram of the total number of steps taken each day after filling the missing values.
```{r, echo=TRUE}
plot(aggregate(steps ~ date, data = act, sum, na.rm = TRUE), type = "h", main = "Total Number of steps per Day")
```

Following is the mean and median values of the total number of steps after filling the missing values.
```{r, echo=TRUE}
mean(act$steps,na.rm = TRUE)
median(act$steps,na.rm = TRUE)
```

And, following is showing each mean and median stpes values per each day after filling the missing values.
```{r, echo=TRUE}
aggregate(steps ~ date, data = act, mean, na.rm = TRUE)
aggregate(steps ~ date, data = act, median, na.rm = TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?

Plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days.
```{r, echo=TRUE}
act <- cbind(act, week = factor(ifelse(weekdays(act$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")))

library(ggplot2)
g <- ggplot(act, aes(interval,steps))
g + geom_line() + facet_grid(week ~ .)

```

